A unified pseudo-$C_\ell$ framework

### 2018-09-18 doc/Paper/main.tex
updated contentious sentence

    While maximum-likelihood~\citep{1998PhRvD..57.2117B,2003PhRvD..67b3001W} or minimum-variance and quadratic estimators~\citep{1997PhRvD..55.5895T,2000MNRAS.317L..23H,2001PhRvD..64f3001T,2018arXiv180702484V} can recover the power spectrum with virtually no loss of information (see also the approaches of e.g. \cite{2004ApJS..155..227E,2008MNRAS.389.1284T,2016MNRAS.455.4452A}, based on Bayesian sampling methods), their computational implementation is prohibitively expensive for high resolution data, since these methods scale as $\ell_{\rm max}^6$ with the maximum multipole $\ell_{\rm max}$ (or $\ell_{\rm max}^4$ for optimal algorithms such as that of \cite{2003PhRvD..67b3001W}). For this reason, pseudo-$C_\ell$ algorithms \citep{1973ApJ...185..413P,2001PhRvD..64h3003W,2001ApJ...561L..11S,2001PhRvD..64h3003W,2002ApJ...567....2H,2002MNRAS.336.1304H,2004MNRAS.350..914C,2005MNRAS.358..833T} have become a popular alternative that preserves the $\ell_{\rm max}^3$ scaling of spherical harmonic transforms (and faster in the flat-sky approximation). This method can be used on fields with arbitrary spin \citep{2003ApJS..148..161K} in both curved and flat skies, has been extended to deal with contaminant deprojection in scalar fields \citep{2017MNRAS.465.1847E}, and can be optimized to measure the CMB $E/B$-mode power spectrum \citep{2002PhRvD..65b3505L,2003PhRvD..67b3501B,2006PhRvD..74h3002S,2009PhRvD..79l3515G,2011PhRvD..83h3003B}. Unfortunately, in spite of the advantages of this method, while a number of power spectrum codes solving some of these problems have been put forward\footnote{See, for instance: \href{http://www2.iap.fr/users/hivon/software/PolSpice/}{PolSpice}, \href{https://gitlab.in2p3.fr/tristram/Xpol}{Xpol}, \href{https://gitlab.in2p3.fr/tristram/Xpure}{Xpure} and \href{https://github.com/thibautlouis/hyperQube}{hyperQube}.}, no public, validated and well-maintained code exists supporting all of these capabilities consistently. This paper presents a public software package, \nmt, that unifies all of these features providing a general framework to estimate pseudo-$C_\ell$ angular power spectra in a wide variety of scenarios. The paper also presents a few novel results that were derived as part of the development of \nmt. These include extending the contaminant deprojection formalism to spin-$2$ fields, combining it with $E/B$ purification and providing a consistent derivation of the same functionality in the flat-sky approximation.

>>>>

    While maximum-likelihood~\citep{1998PhRvD..57.2117B,2003PhRvD..67b3001W} or minimum-variance and quadratic estimators~\citep{1997PhRvD..55.5895T,2000MNRAS.317L..23H,2001PhRvD..64f3001T,2018arXiv180702484V} can recover the power spectrum with virtually no loss of information (see also the approaches of e.g. \cite{2004ApJS..155..227E,2008MNRAS.389.1284T,2016MNRAS.455.4452A}, based on Bayesian sampling methods), their computational implementation is prohibitively expensive for high resolution data, since these methods scale as $\ell_{\rm max}^6$ with the maximum multipole $\ell_{\rm max}$ (or $\ell_{\rm max}^4$ for optimal algorithms such as that of \cite{2003PhRvD..67b3001W}). For this reason, pseudo-$C_\ell$ algorithms \citep{1973ApJ...185..413P,2001PhRvD..64h3003W,2001ApJ...561L..11S,2001PhRvD..64h3003W,2002ApJ...567....2H,2002MNRAS.336.1304H,2004MNRAS.350..914C,2005MNRAS.358..833T} have become a popular alternative that preserves the $\ell_{\rm max}^3$ scaling of spherical harmonic transforms (and faster in the flat-sky approximation). This method can be used on fields with arbitrary spin \citep{2003ApJS..148..161K} in both curved and flat skies, has been extended to deal with contaminant deprojection in scalar fields \citep{2017MNRAS.465.1847E}, and can be optimized to measure the CMB $E/B$-mode power spectrum \citep{2002PhRvD..65b3505L,2003PhRvD..67b3501B,2006PhRvD..74h3002S,2009PhRvD..79l3515G,2011PhRvD..83h3003B}. Due to the advantages of this method, a number of useful, publicly available codes currently exist to carry out some of these calculations \footnote{See, for instance: \href{http://www2.iap.fr/users/hivon/software/PolSpice/}{PolSpice}, \href{https://gitlab.in2p3.fr/tristram/Xpol}{Xpol}, \href{https://gitlab.in2p3.fr/tristram/Xpure}{Xpure} and \href{https://github.com/thibautlouis/hyperQube}{hyperQube}.}. Unfortunately, no public, validated and maintained code exists simultaneously supporting all of the capabilities described above in a consistent manner. This paper presents a public software package, \nmt, that unifies all of these features providing a general framework to estimate pseudo-$C_\ell$ angular power spectra in a wide variety of scenarios. The paper also presents a few novel results that were derived as part of the development of \nmt. These include extending the contaminant deprojection formalism to spin-$2$ fields, combining it with $E/B$ purification and providing a consistent derivation of the same functionality in the flat-sky approximation.
